---
title: "Case Study 1"
author: "Meredith Ludlow & Kito Patterson"
date: "October 9, 2018"
output: 
  html_document:
    keep_md: true
---
## Introduction
Write intro here.

I used the following R packages in the course of this analysis: plyr
```{r echo=TRUE}
library(plyr)
library(dplyr)
library(ggplot2)
```

First, we will look at the number of breweries in each state. After we imported the brewery data, we created a table of the number of breweries that are located in each state. The table is shown below.
```{r echo=TRUE}
Breweries <- read.csv("Breweries.csv", header=TRUE)
BreweriesPerState <- count(Breweries, "State")
BreweriesPerState
# Should we make a histogram?
```

After importing the beer data, we merged the two data sets together on the variable Brew_ID. This variable contains the specific ID numbers for each brewery. The first and last 6 lines of the merged data set are shown below.
```{r echo=TRUE}
Beers <- read.csv("Beers.csv", header=TRUE)
colnames(Beers)[5] <- "Brew_ID"
BrewsandBeers <- merge(Beers, Breweries, by = "Brew_ID")
colnames(BrewsandBeers)[2] <- "BeerName"
colnames(BrewsandBeers)[8] <- "BrewerieName"
head(BrewsandBeers)
tail(BrewsandBeers)
```

For each column in the merged data set, the number of missing data points are summed. The results are listed in the table below.
```{r echo=TRUE}
na_counts <- sapply(BrewsandBeers, function(y) sum(is.na(y)))
na_counts
```

Median alcohol content and international bitterness unit by state
```{r echo=TRUE}
MedianbyState <- group_by(BrewsandBeers, State) %>%
summarise(median_ABV = median(ABV, na.rm=TRUE),
median_IBU = median(IBU, na.rm=TRUE))

MedianbyState
```

State with the most alcohol content 
```{r echo=TRUE}
MaxABVState <- select(BrewsandBeers, State, ABV) %>%
slice(which.max(ABV))

MaxABVState
```

State with the most bitterness
```{r echo=TRUE}
MaxIBUState <- select(BrewsandBeers, State, IBU) %>%
slice(which.max(IBU))

MaxIBUState
```
Descriptive summary statistics
```{r echo=TRUE}
summary(BrewsandBeers$ABV)
```





