---
title: "Case Study 1"
author: "Meredith Ludlow & Kito Patterson"
date: "October 9, 2018"
output: 
  html_document:
    keep_md: true
---
### Introduction
We analyzed two data sets, one containing information about breweries across the US and one containing information about types of beers, to look at the distrubution of breweries per state. We also focused on the international bitterness unit (IBU) and alchohol level to see which states have the highest IBU and alchol content and if there is a correlation between the two variables.
</br>

We used the following R packages in the course of this analysis: plyr, dplyr, stringr and ggplot2.
```{r echo=TRUE}
library(plyr)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
```
</br>


First, we will look at the number of breweries in each state. After we imported the brewery data, we counted the number of breweries that are located in each state. We put the information in a histogram, which is shown below. Colorado and California have the highest number of breweries of all states.
```{r echo=TRUE}
Breweries <- read.csv("Breweries.csv", header=TRUE)
BreweriesPerState <- count(Breweries, State)
BpS <- ggplot(data=BreweriesPerState, aes(x=State, y=n, fill=State)) + 
  geom_bar(stat='identity', colour = 'black') +
  coord_flip() + 
  ggtitle("Number of Breweries per State") + ylab("Frequency") +
  theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5))
BpS
```
</br>


After importing the beer data, we merged the two data sets together on the variable Brew_ID. This variable contains the specific ID numbers for each brewery and is a variable that both data sets have. The first and last 6 lines of the merged data set are shown below to show the merged data set.
```{r echo=TRUE, results='asis'}
Beers <- read.csv("Beers.csv", header=TRUE)
colnames(Beers)[5] <- "Brew_ID"
BrewsandBeers <- merge(Beers, Breweries, by = "Brew_ID")
colnames(BrewsandBeers)[2] <- "BeerName"
colnames(BrewsandBeers)[8] <- "BrewerieName"
kable(head(BrewsandBeers), caption="First 6 Lines")
kable(tail(BrewsandBeers), caption = "Last 6 Lines")
```
</br>


For each variable in the merged data set, the number of missing data points are summed. The results are listed in the table below. Note that IBU has a very high number of missing data points compared to alcohol content (ABV).
```{r echo=TRUE, render="asis"}
na_counts <- sapply(BrewsandBeers, function(y) sum(is.na(y)))
kable(na_counts, caption = "The Number of Missing Data for each Variable")
```
</br>


Next, we group the merged data by state and then find the median alcohol content and bitterness for each state. We don't take into account the missing data points when finding the median. Then the median ABV and IBU for each state are displayed in the histograms below. It seems like the median alcohol content varies less by state than the median bitterness does.
```{r echo=TRUE}
MedianbyState <- group_by(BrewsandBeers, State) %>%
summarise(median_ABV = median(ABV, na.rm=TRUE),
median_IBU = median(IBU, na.rm=TRUE))

bp_ABV <- ggplot(data=MedianbyState, aes(x=State, y=median_ABV, fill=State)) +
  geom_bar(stat='identity', colour = 'black') +
  coord_flip() + 
  ggtitle("Median Alcohol Content by State") + ylab("Median Alcohol Content") +
  theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5))

bp_IBU <- ggplot(data=MedianbyState, aes(x=State, y=median_IBU, fill=State)) +
  geom_bar(stat='identity', colour = 'black') +
  coord_flip() + 
  ggtitle("Median Bitterness Level by State") + ylab("Median Bitterness") +
  theme(legend.position="none") +
  theme(plot.title = element_text(hjust = 0.5))

suppressWarnings(print(bp_ABV))
suppressWarnings(print(bp_IBU))
```
</br>


We calculate the maximum alcohol content and bitterness level out of all of the different beers to see which state has the highest ABV and IBU. Oregan has the higest alcohol content at 0.128 and Oregan has the highest bitterness at 138.
```{r echo=TRUE, results='asis'}
MaxABVState <- select(BrewsandBeers, State, ABV) %>%
slice(which.max(ABV))
MaxIBUState <- select(BrewsandBeers, State, IBU) %>%
slice(which.max(IBU))

kable(MaxABVState)
kable(MaxIBUState)
```
</br>


Below are summary statistics for the alcohol content of all beers in the data set. These statistics include the minimum, maximum, mean, and median, among others.
```{r echo=TRUE}
summary(BrewsandBeers$ABV)
```
</br>


Below is a scatterplot that plots ABV and IBU values. There seems to be a positive correlation between IBU and ABV. As the alcohol content increases, so does the bitterness level. The state of Oregon has the highest recorded bitterness level. This could also mean that Orgeon has one of the highest levels of alcohol content given there seems to be a correlation between bitterness and alcohol levels.
```{r echo=TRUE}
sp <- ggplot(BrewsandBeers, aes(x=ABV, y=IBU)) + 
  geom_point(pch = 21, size = 2, color="red") +
  geom_smooth(method="lm") + 
  ggtitle("Alcohol Content vs Bitterness Level") +
  theme(plot.title = element_text(hjust = 0.5))
  

suppressWarnings(print(sp))
```
</br>

We created an additional variable that captures the beer type from the string values found in the Style column. Ale and IPA are the most popular Beer types followed by APA, Lager and Stout. It would be interesting to better understand what area's of the country contribute to this popularity. Also, there may be more demand for beer styles that aren't being currently produced by local breweries. Deeper analysis is required to answer these questions and others. 
```{r echo=TRUE}
BrewsandBeers$StyleCategory <- word(BrewsandBeers$Style,-1)

StyleCategory_Counts <- BrewsandBeers %>%
group_by(StyleCategory) %>%
summarise(count = n()) %>%
arrange(desc(count))

btp <- ggplot(data=StyleCategory_Counts, aes(x=StyleCategory, y=count, fill=StyleCategory)) + 
geom_bar(stat='identity', colour = 'black') + 
coord_flip() +
ggtitle("Beer Type Counts") + ylab("Frequency") +
theme(legend.position="none") +
theme(plot.title = element_text(hjust = 0.5))

suppressWarnings(print(btp))
kable(head(StyleCategory_Counts,n=10))
```



##Conculsion
In conclusion, the western territories seem to have the highest concentration of breweries led by California, Colorado, Oregon and Washington. It was shown above there seems to be a positive linear relationsip between the alchol content and bitterness level. As alcohol increses so does bitterness. This would suggest that the number of breweries is somewhat driven by the alcohol and bitterness levels in beers produced by these breweries (California and Oregon both have the highest recorded alcohol and bitterness respectively). Further analysis is required to answer other questions of interest like percent of total Beer styles produced by state/city sliced by albohol and bitterness values. 
</br>


Master Branch
https://github.com/MeredithLudlow/DDS-CaseStudy1.git

Forked Master Branch
https://github.com/bbal20/DDS-CaseStudy1.git



